---
title: "mappER - Thiessen Polygons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mappER - Thiessen Polygons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

Currently the `mappER` only has the functionality to support Thiessen (Voronoi) 
polygons. To access this functionality you can load `flode` or mappER directly.

```{r setup}
library(mappER)
```

```{r flode, eval = FALSE}
library(flode)
```

## Loading gauges

The best way to get gauge data into `mappER` is to use the `riskyData` functions
 such as `loadAPI()`.
 
```{r risky, eval = FALSE}
 library(riskyData)
 crowle <- loadAPI(ID = '457592',
                  measure = 'rainfall',
                  period = 900, 
                  type = 'total',
                  datapoints = 'earliest')
```

To save time we will use the pre-loaded data from the package;

```{r loadGauges}
data(crowle); data(bickley); data(barnhurst); data(hollies); data(ledbury); 
data(bettwsYCrwyn)
```

To compile these sites into a spatial dataset we can use the `getCoords()` 
function;

```{r getCoords}
gauges <- getCoords(crowle,
                    bickley,
                    barnhurst,
                    hollies,
                    ledbury,
                    bettwsYCrwyn)
gauges
```

If you wished to plot these data, you can do so using `leaflet`, an open-source 
JavaScript library for interactive maps.

```{r map1}
library(leaflet)
leaflet(gauges) %>%
  addTiles() %>%
  addCircleMarkers(color = "orangered",
                   radius = 6,
                   fillOpacity = 0.8,
                   stroke = "black",
                   label = ~stationName,
                   labelOptions = labelOptions(permanent = TRUE,
                                               noHide = TRUE,
                                               direction = "bottom"))
```

## Developing the Thiessen plots

A Thiessen polygon is a Voronoi Diagram that is also referred to as the 
Dirichlet Tessellation. Given a set of points, it defines a region around each 
point. A Thiessen polygon divides a plane such that each point is enclosed 
within a polygon and assigns the area to a point in the point set. 

In forecasting we use these to determine gauge proportion. With `mappER` they 
are incredibly easy to develop. We can use the `teeSun()` function with the 
above gauge data to develop the tessellation.

```{r teeSun}
voronoi <- teeSun(gauges)
```

```{r plot2, echo = FALSE}
leaflet(gauges) %>%
  addTiles() %>%
  addPolygons(data = voronoi,
              color = "black",
              fillOpacity = 0.1,
              weight = 1.5) %>%
  addCircleMarkers(data = gauges,
                   color = 'orangered',
                   radius = 6,
                   fillOpacity = 0.8,
                   stroke = 'black',
                   label = ~stationName,
                   labelOptions = labelOptions(permanent = TRUE,
                                               noHide = TRUE,
                                               direction = "bottom"))
```

## Intersecting the Thiessen with a catchment

To load a catchment shapefile we can use the `loadCatchment()` funtion.

```{r loadcatch, eval = FALSE}
loadCatchment(filepath = "c://abcd/efg/hi/Catchment.shp")
```

```{r catchment, echo = FALSE}
data("bewdCatch")

leaflet(bewdCatch) %>%
  addTiles() %>%
  addPolygons()
```

Now we have the catchment loaded, it can be intersected by the Thiessen 
polygons. This can be done using the `intersectPoly()` function.

```{r intersect}
int <- intersectPoly(coords = gauges,
                     voronoi = voronoi,
                     catchment = bewdCatch)
```

To plot with a custom colour scheme;

```{r plot3}
colors <- colorFactor(palette = c("blue", "green", "yellow", "red"),
                      domain = (int$stationName))
leaflet(int) %>%
  addTiles() %>%
  addPolygons(color = colors(int$stationName))
```

We can obtain more details on the gauges when we use the `gaugeProp()` function,
 this displays the proportion of catchment that each rain gauge covers. It uses 
 the intersected polygon as the basis for its measurements. 

```{r gaugeProp, eval = FALSE}
gaugeProp(int)
```

```{r gaugeProp2, echo = FALSE}
gp <- gaugeProp(int)

gp %>%  gt::gt()
```

These outputs can then be applied in `riskyData` to generate catchment averaged 
rainfall.

## Putting it all together

To plot all the elements simultaneously we can use the `teeSunPlot()` function;

```{r teeSunPlot}
teeSunPlot(gaugeCoords = gauges, intersectPoly = int)
```

